---
title: "Hands-on Exercise 1"
author: "Lim Li Ying"
---

# Getting Started

## Installing and launching R packages

The code chunk below uses `p_load()` of the `pacman` package to check if the `tidyverse` packages are installed on the computer. If they are, they will then be loaded into the R environment.

```{r}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
pacman::p_load(tidyverse)
```

## Importing the data

To import the data *"Exam_data.csv"* into the R environment, `read_csv()` of the `readr` package is used, as seen in the code chunk below.

```{r}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
exam_data <- read_csv("data/Exam_data.csv")
```

# Exploring ggplot2

## Bar chart

```{r}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
ggplot(data = exam_data, 
       aes(x = RACE)) +
  geom_bar()
```

## Dot plot

```{r}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
ggplot(data = exam_data, aes(x = MATHS)) + 
  geom_dotplot(binwidth = 2.5, dotsize = 0.5) +
  scale_y_continuous(NULL, breaks = NULL)
```

-   `scale_y_continuous` is used to remove the y-axis.

## Histogram

```{R}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram()
```

-   By default, the bin width is 30.

### Modifying the histogram by changing `geom()`

In the code chunk below:

-   *bins* argument is used to change the number of bins to 20

-   *color* argument is used to change the outline color of the histogram to black

-   *fill* argument is used to change the fill color of the histogram to violet

```{R}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, color = "black", fill = "violet")
```

### Modifying the histogram by changing `aes()`

The code chunk below changes the fill color of the histogram using the subgroup of `aes()`, which in this case, is the gender of the students.

```{R}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
ggplot(data = exam_data, aes(x = MATHS, fill = GENDER)) +
  geom_histogram(bins = 20, color = "grey30")
```

## Kernel density estimate

```{R}
#| code-fold: show
#| code-summary: "Click to show/hide the code"
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_density()
```
