---
title: "Take-home Exercise 3"
author: "Lim Li Ying"
date: "22 May 2023"
edit: visual
execute: 
  warning: false
---

# 1 The Task

With reference to [Mini Challenge 3 of the VAST Challenge 2023](https://vast-challenge.github.io/2023/MC3.html), the task is to use visual analytics to identify anomalies in the business groups present in the knowledge graph. 

# 2 The Data

The data set used in this exercise *MC3.json* was obtained from the VAST Challenge 2023 website.


# 3 Data Preparation

## 3.1 Installing and loading of R packages

In this exercise, the following R packages will be used:

1.  `tidyverse`: for data cleaning and manipulation.

2.  `lubridate`:

3.  `jsonlite`: for loading and reading of the *.json* file.

4.  `visNetwork`: for creating interactive network graphs.

The code chunk below uses `p_load()` of the *pacman* package to check if all the aforementioned packages are installed, and install the packages are yet to be installed. The packages are then loaded into the R environment.

```{r}
pacman::p_load(igraph, tidygraph, ggraph, visNetwork, lubridate, clock, tidyverse, graphlayouts, jsonlite)
```

## 3.2 Importing and loading the data set

To import the data *"mc2_challenge_graph.json"* file into the R environment, `fromJSON()` of the *jsonlite* package is used, as seen in the code chunk below.

```{r}
MC3 <- jsonlite::fromJSON("data/MC3.json")
```

The `glimpse()` function from the *dpylr* package is used to see a general overview of the data set.

```{r}
glimpse(MC3)
```

```{r}
MC3_edges <- as_tibble(MC3$links)
```

```{r}
MC3_edges
```
